---
title: "Evidence based decision making"
subtitle: "Learn how to understand and apply data"
format:
  html:
    css: include/webex.css
    include-after-body: include/webex.js
---

```{r, echo = FALSE, results='asis'}
knitr::opts_chunk$set(echo = FALSE)
library(webexercises)
# Uncomment to change widget colours:
#style_widgets(incorrect = "goldenrod", correct = "purple", highlight = "firebrick")
```

## Example Questions

### Fill-In-The-Blanks (`fitb()`)

-   2 + 2 is `r fitb(4)`

```{r}
x <- sample(2:8, 1)
```

-   The square root of `r x^2` is: `r fitb(x)`
-   What is the letter after D? `r fitb("E", ignore_case = TRUE)`
-   How do you load the tidyverse package? `r fitb(c("library( tidyverse )", "library( \"tidyverse\" )", "library( 'tidyverse' )"), ignore_ws = TRUE, width = "20")`
-   Type a vowel: `r fitb(c("A", "E", "I", "O" , "U"), ignore_case = TRUE)`
-   Type any 3 letters: `r fitb("^[a-zA-Z]{3}$", width = 3, regex = TRUE)`

### Multiple Choice (`mcq()`)

-   "Never gonna give you up, never gonna: `r mcq(c("let you go", "turn you down", "run away", answer = "let you down"))`"
-   "I `r mcq(c(answer = "bless the rains", "guess it rains", "sense the rain"))` down in Africa" -Toto

### True or False (`torf()`)

-   True or False? You can permute values in a vector using `sample()`. `r torf(TRUE)`

### Longer MCQs (`longmcq()`)

**What is a p-value?**

```{r}
opts_p <- c(
   "the probability that the null hypothesis is true",
   answer = "the probability of the observed, or more extreme, data, under the assumption that the null-hypothesis is true",
   "the probability of making an error in your conclusion"
)
```

`r longmcq(opts_p)`

**What is true about a 95% confidence interval of the mean?**

```{r}
# use sample() to randomise the order
opts_ci <- sample(c(
  answer = "if you repeated the process many times, 95% of intervals calculated in this way contain the true mean",
  "there is a 95% probability that the true mean lies within this range",
  "95% of the data fall within this range"
))
```

`r longmcq(opts_ci)`

## Checked sections

::: {.webex-check .webex-box}
I am going to learn a lot: `r torf(TRUE)`

```{r, results='asis'}
opts <- c(
   "the probability that the null hypothesis is true",
   answer = "the probability of the observed, or more extreme, data, under the assumption that the null-hypothesis is true",
   "the probability of making an error in your conclusion"
)

cat("What is a p-value?", longmcq(opts))
```
:::

## Hidden solutions and hints

```{r}
with(cars, plot(speed, dist))
```

`r hide("I need a hint")`

See the documentation for `plot()` (`?plot`)

`r unhide()`

<!-- note: you could also just set webex.hide to TRUE -->

```{r, echo = TRUE, eval = FALSE}
#| webex.hide: "Click here to see the solution"
plot(cars$speed, cars$dist)
```
